---
title: "p8105_hw5_bac2214"
author: "Brianna Carnagie"
date: "2023-11-12"
output: github_document
---

### Question 1

```{r setup, include=FALSE}

library(tidyverse)
library(dplyr)
library(purrr)
library(broom)
library(ggplot2)
library(readr)
library(stringr)
library(viridis)

homicide_df = read_csv("data/homicide-data.csv") |> 
  mutate(
    city_state = paste(city, state, sep = ",")
    )

homicides_by_city = read_csv("data/homicide-data.csv") |> 
  mutate(
    city_state = paste(city, state, sep = ",")) |> 
  group_by(city_state) |> 
  summarize(total_homicides = n(),
            unsolved_homicides = sum(disposition %in% c("Closed without arrest", "Open/No arrest")))

baltimore_data = homicides_by_city |> 
  filter(city_state == "Baltimore,MD") 
  
baltimore_prop_test = 
  prop.test(x = baltimore_data$unsolved_homicides,
            n = baltimore_data$total_homicides) |> 
  broom::tidy() |> 
  mutate(
    est_prop = estimate[1],
    conf_int = paste(conf.low, conf.high, sep = ",")) |> 
  select(est_prop,conf_int)
``` 
The raw data contains details about the victim of a homicide including their name, race, age, and sex. There are also details regarding the location down to the latitude and longitude as well as the disposition of the person who committed the homicide.


I am now repeating the steps for each city and state value in the homicides_by_city dataframe.
```{r}

df_to_list = split(homicides_by_city, seq(nrow(homicides_by_city)))

prop_and_ci = function(data) {
  result = prop.test(x = data$unsolved_homicides, n = data$total_homicides)
  tidy_result = tidy(result)
  estimated_proportion = tidy_result$estimate[1]
  conf_low = tidy_result$conf.low
  conf_high = tidy_result$conf.high
  return(data.frame(estimated_proportion, conf_low, conf_high))
}

output = map(df_to_list, prop_and_ci) |> 
  bind_rows() |> 
  mutate(city_state = unique(homicides_by_city$city_state)) |> 
  select(city_state,estimated_proportion:conf_high)

```

Here is the plot of estimates and confidence intervals for unsolved homicides by city.

```{r}

ggplot(output, aes(x = reorder(city_state, -estimated_proportion), y = estimated_proportion)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = conf_low, ymax = conf_high), width = 0.2) +
  coord_flip() +  # Flip the x and y axes for horizontal bars
  labs(x = "City State", y = "Estimated Proportion") +
  theme_minimal() +  # Apply a minimal theme
  ggtitle("Estimates and Confidence Intervals for Unsolved Homicides by City") +
  theme(plot.title = element_text(hjust = 0.5))  # Center the title
```

The plot shows that Tulsa has the lowest proportion of unsolved homicides while Chicago has the highest proportion of unsolved homicides.


### Question 2
```{r}

file_names = list.files(path = "data/question 2", pattern = ".csv", full.names = TRUE)

tidy_study_data = lapply(file_names, function(file) {
  
  df = read.csv(file)  
  subject_id = str_extract(basename(file), "(con|exp)_\\d+")
  arm = ifelse(grepl("con_", basename(file)), "control", "experimental")
  
  df = df |> 
    pivot_longer(cols = c(week_1, week_2, week_3, week_4, week_5, week_6, week_7, week_8), 
                 names_to = "week", values_to = "observation") |> 
    mutate(
      subject = subject_id,
      arm = arm,
      week = as.numeric(gsub("week_", "", week))
    )
  return(df)
  
}) |> bind_rows() |> select(week, arm, subject, observation)

study_data_plot = ggplot(tidy_study_data, aes(x = week, y = observation, group = subject, color = arm)) +
  geom_line() +
  labs(x = "Week", y = "Observation") +
  ggtitle("Study Observations by Week") 

```

The observations for the experimental group tend to be higher over the weeks compared to the control group. However, there is a considerable variation within both groups, and some subjects in the control group have observations that overlap with those in the experimental group.
